Psychological Needs During Intergroup Contact

### 2021-10-25 Methods-and-Results.Rmd
effect size trials

# divide into trait and state
dtWorkerSupp$workerWithinBetween <-
  MlTraitState(
    data = dtWorkerSupp$workerInteractionType,
    id = "PID",
    selection =
      c(
        "keyMotiveFulfilled",
        "competence.daily.all",
        "autonomy.daily.all",
        "relatedness.daily.all",
        "thermometerDutch_1",
        "keymotive_fulfillemt_1",
        "competence_1",
        "autonomy_1",
        "relatedness_1"
      )
  )
```

@@

\paragraph{Core Need}t be explained by whether or not the interaction fulfilled the person's core situational need. This should, in turn, be due to a higher perceived interaction quality. We will this sequentially test whether the fulfillment of the core need during an interaction is (1) related to more positive outgroup attitudes, (2) higher perceived interaction quality, and (3) whether the variance explained by the core need is assumed by the perceived interaction quality if considered jointly.

>>>>

# divide into trait and state
dtWorkerSupp$workerWithinBetween <-
  MlTraitState(
    data = dtWorkerSupp$workerInteractionType,
    id = "PID",
    selection =
      c(
        "keyMotiveFulfilled",
        "competence.daily.all",
        "autonomy.daily.all",
        "relatedness.daily.all",
        "thermometerDutch_1",
        "keymotive_fulfillemt_1",
        "competence_1",
        "autonomy_1",
        "relatedness_1"
      )
  )
dtWorkerSupp$workerOutWithinBetween <-
  MlTraitState(
    data = dtWorkerSupp$workerInteractionType %>% filter(OutgroupInteraction == "Yes"),
    id = "PID",
    selection =
      c(
        "keyMotiveFulfilled",
        "competence.daily.all",
        "autonomy.daily.all",
        "relatedness.daily.all",
        "thermometerDutch_1",
        "keymotive_fulfillemt_1",
        "competence_1",
        "autonomy_1",
        "relatedness_1"
      )
  )

```

@@

\paragraph{Core Need}The main proposal of our article is that the success of an outgroup interaction might be explained by whether or not the interaction fulfilled the person's core situational need. This should, in turn, be due to a higher perceived interaction quality. We will this sequentially test whether the fulfillment of the core need during an interaction is (1) related to more positive outgroup attitudes, (2) higher perceived interaction quality, and (3) whether the variance explained by the core need is assumed by the perceived interaction quality if considered jointly.

```{r workerOutModelMLAttNull, include=FALSE}
# create empty list to organize models
mdlWorkerOut <- 
  list(
    Att = list(),
    Qlt = list()
  )

# Create and save Model
mdlWorkerOut$Att$lmerNull <-
  lme4::lmer(thermometerDutch_1 ~ 1 + (1 | PID), 
             data = dtWorkerSupp$workerOutgroupInteraction) # use optim if it does not converge
mdlWorkerOut$Att$lmeNull <-
  lme(
    thermometerDutch_1 ~ 1,
    random = ~ 1 | PID,
    data = dtWorkerSupp$workerOutgroupInteraction,
    control = list(opt = "nlmimb")
  ) # use optim if it does not converge

# Get summary with p-values (Satterthwaite's method)
# summary(Null.Out.ML.r) #or with the lme function
summ(mdlWorkerOut$Att$lmerNull, digits = 3, center = TRUE)

# Save variances
mdlWorkerOut$Att$varNull <- 
  VarCorr(mdlWorkerOut$Att$lmeNull) # save variances
# The estimate of (between-group or Intercept variance, tau_{00}^2):
mdlWorkerOut$Att$tauNull <- 
  as.numeric(mdlWorkerOut$Att$varNull[1])
# and the estimate of (within-group or residual variance, sigma^2) is:
mdlWorkerOut$Att$sigmaNull <- 
  as.numeric(mdlWorkerOut$Att$varNull[2])
# The ICC estimate (between/between+within) is:
mdlWorkerOut$Att$IccNull <-
  (as.numeric(mdlWorkerOut$Att$varNull[1]) / (as.numeric(mdlWorkerOut$Att$varNull[1]) + as.numeric(mdlWorkerOut$Att$varNull[2])))
mdlWorkerOut$Att$IccPercNull <-
  ((as.numeric(mdlWorkerOut$Att$varNull[1]) / (as.numeric(mdlWorkerOut$Att$varNull[1]) + as.numeric(mdlWorkerOut$Att$varNull[2])))) * 100
```
```{r workerOutModelInterceptAttCore, include=FALSE}
# Create and save Model
mdlWorkerOut$Att$lmeInterceptCore <-
  lme(
    thermometerDutch_1 ~ keymotive_fulfillemt_1_cwc,
    random = ~ 1 | PID,
    data = dtWorkerSupp$workerOutWithinBetween
  )

# Get summary with p-values (Satterthwaite's method)
summ(
  mdlWorkerOut$Att$lmerInterceptCore <-
    lmer(
      thermometerDutch_1 ~ keymotive_fulfillemt_1_cwc + (1 | PID),
      data = dtWorkerSupp$workerOutWithinBetween
    ),
  confint = TRUE,
  digits = 3,
  center = FALSE
)

# Generate 95% parametric bootstrap CIs (and save them as a csv-file):
# write.csv(confint(lmer(thermometerDutch_1 ~ Contact_dum + inNonDutch + (1|PID),data=workerOutgroupInteraction),
#                  method="boot",nsim=1000, parallel = "multicore",
#                  ncpus = 4, seed = 42),
#          "output/tables/ML-Inter-CI.csv")

# Compare new model to previous step
anova(mdlWorkerOut$Att$lmeNull, 
      mdlWorkerOut$Att$lmeInterceptCore) %>%
  as.data.frame() %>%
  select(-call) %>%
  mutate(
    L.Ratio = round(L.Ratio, 3),
    `p-value` = ifelse(`p-value`>=.001, round(`p-value`, 3), "< .001")
  ) %>%
  replace(is.na(.), "") %>%
  kbl(
    .,
    caption = "Worker: Model Comparison",
    format = "html",
    linesep = "",
    booktabs = T,
    align = rep("c", ncol(.)),
    digits = 3
  ) %>%
  kable_styling(position = "left")

# Save variances
mdlWorkerOut$Att$varInterceptCore <-
  lme4::VarCorr(mdlWorkerOut$Att$lmeInterceptCore)

# The estimate of between-group (or Intercept variance) explained:
# Variance Explained = 1 – (Var with Predictor/Var without Predictor)
mdlWorkerOut$Att$varBtwInterceptCore <-
  1 - (mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["(Intercept)", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varBtwPercInterceptCore <-
  (1 - (mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["(Intercept)", "Variance"] %>% as.numeric)) * 100
# and the estimate of within-group (or residual variance) explained is:
mdlWorkerOut$Att$varWithinInterceptCore <-
  1 - (mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["Residual", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varWithinPercInterceptCore <-
  (1 - (mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varNull["Residual", "Variance"] %>% as.numeric)) * 100
```
```{r workerOutModelSlopesAttCore, include=FALSE}
# Create and save Model (optimizer needed to reach convergence)
mdlWorkerOut$Att$lmeSlopesCore <-
  lme(
    thermometerDutch_1 ~
      keymotive_fulfillemt_1_cwc,
    random = ~ 1 + keymotive_fulfillemt_1_cwc | PID,
    control = lmeControl(opt = "optim"),
    data = dtWorkerSupp$workerOutWithinBetween
  )

# Get summary with p-values (Satterthwaite's method) [+ save model for plotting]
summ(
  mdlWorkerOut$Att$lmerSlopesCore <- lmer(
    thermometerDutch_1 ~
      keymotive_fulfillemt_1_cwc +
      (1 + keymotive_fulfillemt_1_cwc | PID),
    data = dtWorkerSupp$workerOutWithinBetween
  ),
  confint = TRUE,
  digits = 3,
  center = FALSE
)

mdlWorkerOut$Att$lmerSlopesCoreCI <- confint(mdlWorkerOut$Att$lmerSlopesCore)

r2mlm::r2mlm(mdlWorkerOut$Att$lmerSlopesCore, bargraph = TRUE)
mitml::multilevelR2(mdlWorkerOut$Att$lmerSlopesCore)
sjstats::r2(mdlWorkerOut$Att$lmerSlopesCore, mdlWorkerOut$Att$lmerNull)
performance::r2(mdlWorkerOut$Att$lmerSlopesCore)
performance::model_performance(mdlWorkerOut$Att$lmerSlopesCore)
performance::compare_performance(mdlWorkerOut$Att$lmerNull, mdlWorkerOut$Att$lmerSlopesCore)

# Generate 95% parametric bootstrap CIs (and save them as a csv-file):
# write.csv(confint(model.ran0,
#               method="boot",nsim=1000,
#               parallel = "multicore", ncpus = 4, seed = 42),
#          "output/tables/ML-RandomSlopes-CI.csv")

# Compare new model to previous step
anova(mdlWorkerOut$Att$lmeNull, 
      mdlWorkerOut$Att$lmeInterceptCore, 
      mdlWorkerOut$Att$lmeSlopesCore) %>%
  as.data.frame() %>%
  select(-call) %>%
  mutate(
    L.Ratio = round(L.Ratio, 3),
    `p-value` = ifelse(`p-value`>=.001, round(`p-value`, 3), "< .001")
  ) %>%
  replace(is.na(.), "") %>%
  kbl(
    .,
    caption = "Worker: Model Comparison",
    format = "html",
    linesep = "",
    booktabs = T,
    align = rep("c", ncol(.)),
    digits = 3
  ) %>%
  kable_styling(position = "left")

# Save variances
mdlWorkerOut$Att$varSlopesCore <- 
  lme4::VarCorr(mdlWorkerOut$Att$lmerSlopesCore)

# The estimate of between-group (or Intercept variance) explained:
# Variance Explained = 1 – (Var with Predictor/Var without Predictor)
mdlWorkerOut$Att$varBtwSlopesCore <-
  1 - (mdlWorkerOut$Att$varSlopesCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varBtwPercSlopesCore <-
  (1 - (mdlWorkerOut$Att$varSlopesCore["(Intercept)", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["(Intercept)", "Variance"] %>% as.numeric)) * 100
# and the estimate of within-group (or residual variance) explained is:
mdlWorkerOut$Att$varWithinSlopesCore <-
  1 - (mdlWorkerOut$Att$varSlopesCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric)
mdlWorkerOut$Att$varWithinPercSlopesCore <-
  (1 - (mdlWorkerOut$Att$varSlopesCore["Residual", "Variance"] %>% as.numeric / mdlWorkerOut$Att$varInterceptCore["Residual", "Variance"] %>% as.numeric)) * 100

# Assumption Checks:
mdlWorkerOut$Att$diagSlopesCore <- 
  sjPlot::plot_model(mdlWorkerOut$Att$lmerSlopesCore, type = "diag")
grid.arrange(
  mdlWorkerOut$Att$diagSlopesCore[[1]],
  mdlWorkerOut$Att$diagSlopesCore[[2]]$`PID`,
  mdlWorkerOut$Att$diagSlopesCore[[3]],
  mdlWorkerOut$Att$diagSlopesCore[[4]]
)

# Plot prediction model
mdlWorkerOut$Att$predSlopesCore <- 
  dtWorkerSupp$workerOutWithinBetween %>%
  select(thermometerDutch_1, session, PID) %>% 
  mutate(measure = predict(mdlWorkerOut$Att$lmeSlopesCore,
                           dtWorkerSupp$workerOutWithinBetween,
                           re.form = NA
                           )
         )

(
  mdlWorkerOut$Att$predPltSlopesCore <-
    ggplot(data = mdlWorkerOut$Att$predSlopesCore, aes(x = session, y = measure)) +
    geom_line(alpha = 1, color = "blue") +
    geom_line(aes(y = thermometerDutch_1), alpha = 1) +
    facet_wrap( ~ PID, ncol = 6) +
    xlab("Time") +
    ylab("Outgroup Attitudes") +
    theme_Publication()
)
ggsave(
  filename = "Figures/WorkerOut_PredictionPlot_SlopesAttCore.png",
  mdlWorkerOut$Att$predPltSlopesCore,
  width = 18,
  height = 12,
  dpi = 800,
  units = "cm",
  device = "png"
)
```
